---
title: "Q1"
author: "Arindom Baruah (32779267)"
format: html
execute: 
  eval: true
  warning: false
  message: false
---

```{r echo=FALSE}
library(tidyverse)
library(ggplot2)
```


## Exercise 1

Data from a recent study of viewports can be found in the data set `viewports.csv`. The purpose of the data collection was to *you simply do not know how users are going to visit your website or web app. Instead of making design decisions on strict, limited breakpoints, keep in mind the sheer amount of fragmentation there is in viewports.* The data was collected primarily from the USA and Europe, on 122,321 devices. Below is a glimpse of the data. It has three columns defining the width and height of the screen, and count is the number of devices with these dimensions. 

```{r}
#| echo: false
#| eval: true
#| message: false

v <- read_csv(here::here("data/viewports.csv"))
glimpse(v)
```
:::{.callout-note}
# Question A
__How many variables in this data and what type are they?__

:::

Answer a

We observe a total of __`r v %>% ncol()`__ variables. Each of these variables are of the "Double" datatype which implies that they __are numerical values and are continuous in nature__.

:::{.callout-note}
# Question B
__Based on the description above, and the variables available in the data, posit two open-ended questions that might be interesting to answer with the data.__
:::

Answer b

Based on the description above, following are the open-ended questions which may be interesting to answer:

<div class="alert alert-block alert-warning">

1. Does the data suggest if users prefer a relatively narrow devices (such as smartphones) and whether wide devices (such as tablets) are abundant enough among users to account for any necessary website dimensional changes ?
2. Is there a popular dimensional ratio of devices (Width:Height) which users prefer ?
3. Is there a need to create an application which standardises the viewport for the device in which, the application has been installed ?

</div>
:::{.callout-note}
# Question C

__Create a bubble scatterplot, where size represents the count, with an `x=y` guideline underlying the points. Describe what you learn about viewport sizes from this plot.__
:::

Answer c


:::{.callout-note}
# Question D
__Make a transformation of the data to create a new variable, `aspect_ratio`, based on Width and Height, that will allow you to explain the sizes in one of the conventional terms eg 16:9. (Note: use width:height definition of aspect ratio.)__
:::

:::{.callout-note}
# Question E
__Show the aspect ratio as a density plot, where the x axis labels correspond to particular ratios, "1:3", "9:16", "1:1", "16:9", "3:1". (Note: Remember that the data is aggregated already, so that Count needs to the incorporated into computing the density. Setting the axis limits to between 0 and 4 is helpful, too.) Describe the distribution and explain what you learn.__
:::

:::{.callout-note}
# Question F
__What is the most frequent aspect ratio? What (type of) device does this correspond to? (Note: You'll need to do some web searching.)__
:::
