---
title: "Q3"
author: "Arindom Baruah (32779267)"
format: html
execute: 
  warning: false
  message: false
---
```{r echo=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggthemes)
library(ggExtra)
library(plotly)
```


```{r echo=FALSE}
df_tb <- read_csv("data/TB_notifications_2023-08-21.csv")
```
## Exercise 3

The World Health Organisation (WHO) routinely reports statistics on tuberculosis. WHO says: "Every year, 10 million people fall ill with tuberculosis (TB). Despite being a preventable and curable disease, 1.5 million people die from TB each year – making it the world’s top infectious killer." This question is about the statistics reported by WHO on TB notifications. 

The data and dictionary were downloaded from https://www.who.int/teams/global-tuberculosis-programme/data. We have made them available in the template repo.

:::{.callout-note}
# Question A
How many countries are recorded?
:::
There are a total of __`r length(unique(df_tb$country,na.rm=TRUE))`__ countries recorded in the dataset.

:::{.callout-note}
# Question B
What is the earliest and latest years in the data?
:::
The earliest year of the study is __`r min(df_tb$year)`__ and the latest year of study is __`r max(df_tb$year)`__.

:::{.callout-note}
# Question C
There are numerous ways that TB is detected, and this is reflected in the data as `new_sp`, `new_sn`, `new_su`, `new_ep` and `new_oth`. Create a new variable which sums these values for each year, for each country. Make an appropriate plot to show the counts for Australia over time. Describe the pattern, and try to explain it, which might require searching for more information about TB reporting.
:::
<div class="alert alert-block alert-warning">
The terms "new_sp," "new_sn," "new_su," "new_ep," and "new_oth" are commonly used abbreviations and codes in the context of tuberculosis (TB) detection and reporting. They are part of the World Health Organization's (WHO) TB case notification system, which categorizes TB cases into different types based on clinical and epidemiological characteristics. These categories help in tracking and managing TB cases effectively. Here's what each of these terms typically represents:

1. **New Smear-Positive (New SP)**:
   - New SP refers to newly diagnosed TB cases where sputum smear microscopy (a laboratory test) has confirmed the presence of acid-fast bacilli (AFB), which are indicative of active TB disease.
   - "Smear-positive" means that TB bacteria are present in the patient's sputum and can be detected under a microscope. This category is important because smear-positive cases are more infectious and require immediate treatment to reduce transmission.

2. **New Smear-Negative (New SN)**:
   - New SN refers to newly diagnosed TB cases where sputum smear microscopy did not detect AFB, but the patient's clinical symptoms and other diagnostic methods (e.g., culture or molecular tests) suggest active TB disease.
   - Some TB cases may not have detectable AFB in their sputum, especially in the early stages of the disease. These cases are still considered infectious and are treated accordingly.

3. **New Extrapulmonary (New EP)**:
   - New EP refers to newly diagnosed TB cases where the disease is located outside the lungs (extrapulmonary). TB can affect various organs and tissues in the body, and extrapulmonary cases occur when TB is found in places other than the respiratory system.
   - Common sites of extrapulmonary TB include lymph nodes, bones, joints, and the central nervous system.

4. **New Relapse (New SU)**:
   - New SU refers to cases in which a patient who was previously treated for TB and considered cured or successfully treated experiences a recurrence of TB. This is often called a relapse.
   - Relapse cases indicate that the initial treatment may not have been entirely successful or that the patient has been reinfected with TB.

5. **New Other (New OTH)**:
   - New OTH represents other newly diagnosed TB cases that do not fit into the specific categories mentioned above. It is a catch-all category for cases that do not fall into the new SP, new SN, new EP, or new SU categories.

</div>

```{r}
df_grouped <- df_tb  %>% mutate(across(new_sp:new_oth, ~replace_na(., 0))) %>% group_by(year,country) %>% mutate(Metric = new_sp + new_sn + new_su + new_ep + new_oth)
df_aus <- df_grouped %>% filter(country=="Australia")
```

```{r}
#| label: fig-tbaus
#| fig-cap: "New TB cases in Australia"
pl1 <- ggplot(data =df_aus,aes(x =year,
                               y = Metric)) + geom_line() + geom_point(color="blue") + theme_classic() + labs(x = "Year",
                                                                 y = "Total new TB cases detected",caption = "Source: WHO")  + ggtitle("New TB cases in Australia") + theme(plot.title = element_text(hjust = 0.5))

ggplotly(pl1)

```


Based on the plot of new TB cases for the country of Australia in @fig-tbaus, we observe that between the time period of 1980-2001, the first spike in the new TB cases detected were in the year of 1993 but died down for the next couple of years. There was however a resurgence of TB cases observed in the year of 1997 and it continued to grow until the end of 2012 to a total number of 1259 cases per year. However, an encouraging drop in cases from the year of 2013 to 2020 with cases ranging between 480 to  609 cases within that period.

Based on a study by the Australian National University which can be referred to [here](https://nceph.anu.edu.au/research/projects/australian-tuberculosis-risk-factor-survey#:~:text=Australia%20reports%20approximately%201300%20cases,has%20been%20observed%20since%202003.), majority of the TB cases were observed to affect immigrants from nations with poor health care and high pollution. This could be a reason as to why, new TB cases were observed to spike during the early parts of 2000s as more number of immigrants started making their way into Australia. 

:::{.callout-note}
# Question D
Subset the data to Australia and five of it's nearest neighbours (Fiji, Indonesia, New Zealand, Papua New Guinea, and East Timor). Make an appropriate plot to compare the incidence across countries. 
:::
```{r}
#| label: fig-ausnei
#| fig-cap: "Number of TB cases in and around Australia"
options(scipen=1000)
countries <- c("Australia","New Zealand","Papua New Guinea","Indonesia","Timor-Leste","Fiji")

pl2 <- ggplot(data = df_grouped %>% filter(country %in% countries),aes(x =year,
                               y = Metric)) + geom_line() + geom_point(color='darkred')+ theme_classic() + labs(x = "Year",
                                                                 y = "Total new TB cases detected",caption = "Source: WHO")  + theme(plot.title = element_text(hjust = 0.5)) + facet_wrap(~country,scales = "free_y")
ggplotly(pl2)
```

@fig-ausnei illustrates the incidences of TB cases in Australia along with its 5 nearest neighbors which are Fiji, Indonesia, New Zealand, Papua New Guinea and Timor-Leste.

Some important observations are as follows:
<div class="alert alert-block alert-warning">

1. The magnitude of incidences are significantly different from Australia. The countries of Indonesia, Papua New Guinea and Timor-Leste have all reported higher cases than Australia. 
2. On the other hand, New Zealand and Fiji have fewer incidences of TB cases when compared to Australia.

</div>
:::{.callout-note}
# Question E
What is wrong with directly comparing the incidence? What information is missing in order to make the comparison. Without adding new data, come up with a transformation of counts that will fix the problem. Make a new plot, and describe how TB is occurring similarly or differently in Australia and it's nearest neighbours.
:::
Answer E

Based on our analysis in the previous section, we can observe the following issues:
<div class="alert alert-block alert-warning">
1. The number of incidences are quite varied for the 6 countries. While on the face of it, it might look like countries with less incidences are doing better, it maybe a misleading statistic to provide any conclusions.

2. The major issue we can observe in @fig-ausnei is the fact that the plot does not give us an idea of the affected population in the various countries. Each country has __a different size of population__ and it is expected for countries with larger population to have higher incidences. Hence, comparing the incidences may not be the right metric to focus on.

3. Instead of using the incidence (count of new cases per year), we can transform our statistic to prevalence (total count of cases per population of the country).

4. We can use the population statistics provided by The Worldbank which can be accessed [here](https://data.worldbank.org/indicator/SP.POP.TOTL)
</div>
```{r}
df_pop <- read_csv("data/world_population.csv",skip=3)
df_pop_long <- df_pop %>% pivot_longer(cols = "1960":"2022",names_to = "year",values_to = "Population") %>% select(c("Country Name","year","Population"))
df_pop_long$year <- as.double(df_pop_long$year)
df_pop_long <- df_pop_long %>% rename("country" = "Country Name")
df_grouped <- left_join(df_grouped,df_pop_long)

```

```{r}
#| label: fig-perc
#| fig-cap: "Percentage of population affected by TB"


df_neighbors <- df_grouped %>% filter(country %in% countries)
df_neighbors <- df_neighbors %>% group_by(country) %>% mutate("Total_cases" = cumsum(Metric)) %>% select(c("country","year","Total_cases","Population"))



pl4 <- ggplot(data = df_neighbors,aes(x =year,
                               y = 100*Total_cases/Population)) + geom_line() + geom_point(color='darkred')+ 
  theme_classic() + labs(x = "Year",
                         y = "Percentage of affected population",
                         caption = "Source: WHO")  + theme(plot.title = element_text(hjust = 0.5)) + 
  facet_wrap(~country) + ggtitle("Percentage of population affected by TB") 
ggplotly(pl4)
```

Upon successful transformation from incidence to prevalence as illustrated by @fig-perc, we can make the following observations:
<div class="alert alert-block alert-warning">

1. The countries of Papua New Guinea, Timor-Leste and Indonesia are significantly affected by the occurence of TB with a sizeable percentage of population who have been reported to suffer from this disease. These countries are closely followed by Fiji. One of the causes for the prevalence of such a disease could be due to lack of healthcare resources.
2. On the other hand, a very small population of Australia and New Zealand have been observed to be affected by the TB disease. These are relatively wealthier economies and have a robust healtcare system which has been successful to curb the TB disease to a very small population.

</div>
